---
title: "exercise one"
format: pdf
editor: visual
---

```{r}
library(tidyverse)
```

```{r}
anscombe_quartet = readRDS("exercise-set_1-20250213/anscombe quartet.rds")
str(anscombe_quartet)
```

*What does the function str() do?*

It gives you an overview of the data's structure, like the variables and the nature of the variables.

```{r}
anscombe_quartet %>% 
  group_by(dataset) %>% 
  summarise(
    mean_x    = mean(x),
    mean_y    = mean(y),
    min_x     = min(x),
    min_y     = min(y),
    max_x     = max(x),
    max_y     = max(y),
    crrltn    = cor(x, y)
  )

```

*What do the summary statistics tell us about the different datasets?*

They seem virtually identical, with the min and max digits being very similar and means being exactly the same.

```{r}
library(ggplot2)
anscombe_quartet <- ggplot(anscombe_quartet, aes(x=x,y=y)) +
  geom_point() + 
  geom_smooth(method = "lm",formula = "y ~ x") +
  facet_wrap(~dataset)
print(anscombe_quartet)
```

```{r}
ggsave("anscombe_plot.png", plot = anscombe_quartet, width = 5, height = 5, units = "in", dpi = 300)
```

*What do the plots tell us about the different datasets?*

Dataset 1 follows a rough linear trend, no extreme outliers

Dataset 2 is curved, does not follow line of best fit

Dataset 3 roughly linear as well, one extreme outlier

Dataset 4 shows a cluster around one x digit with one extreme outlier

*Describe the relationship between x and y in the different datasets.*

1 As x increases, y increases in a linear fashion - positive relationship

2 Y increases faster when X increases

3 Nearly linear, may be a few outliers that pull the line up or down (positive)

4 None, but outliers create positive slope on fitted line

*Would linear regression be an appropriate statistical model to analyse the x-y relationship in each dataset?*

1 yes, simple straight line

2 no, curved pattern, not linear

3 maybe?

4 No, linear regression would not be representative of cluster pattern

*What conclusions can you draw for the plots and summary statistics?*

Summary statistics can be misleading and mask important aspects of a data set. Pairing them with plots is essential in order to assess any nuances in the data.

### Problem 2

```{r}
datasaurus_dozen = readRDS("exercise-set_1-20250213/datasaurus_dozen.rds")
library(tidyverse)

```

```{r}
str(datasaurus_dozen)
datasaurus_dozen
```

*Print descriptive statistics and make a nicely formatted table*

```{r}
datasaurus_summary <- datasaurus_dozen %>%
  group_by(dataset) %>%
  summarise_if(is.double, list(mean = mean, sd = sd))
print(datasaurus_summary)
```

*Calculate the correlations for x and y*

```{r}
r1 <- lm(x ~ y, datasaurus_dozen)
summary(r1)
```

*Plot their relationships including the line of best fit.*

```{r}
library(ggplot2)
library(GGally)
datasaurus_plot <- ggplot(datasaurus_dozen, aes(x = x, y = y)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "gray") +
  facet_wrap(~ dataset) +
  labs(
    title = "Datasaurus Dozen: Relationship between x and y",
    x = "x",
    y = "y"
  ) +
  theme_minimal()
print(datasaurus_plot)
ggsave("datasaurus_plot.png", plot = datasaurus_plot, width = 5, height = 5, units = "in", dpi = 300)
```
